---
title: "Portfolio 4"
output: html_notebook
---
#### Mini-Exercise #1 

Answer the questions about this peninsula example. Then do your own example from Latin America:
1. Where in the world is the peninsula mapped above?
Spain
2. Is altitude a climate variable? How do the peninsula maps plotted above support your answer?
Yes altitude is a climate variable. The peninsula exemplifies this because different alltitudes have unique climates. 
3. Make a South America map with occurrence data for Jungia pauciflora & J. polita and:
          a. Crop the map to allow discerning more detail in the two ranges
          b. Also include altitude data or climate data, with occurrence points overlayed.
```{r}
SAmerica.worldalt <- crop(all.worldalt, extent(-75, -50, -30, -15))
plot(SAmerica.worldalt[["alt"]])
points(cbind(J_pauciflora$longitude, J_pauciflora$latitude), col="red", pch=20, cex=.8)
points(cbind(J_paniculata$longitude, J_paniculata$latitude), col="blue", pch=20, cex=.8)
points(cbind(J_polita$longitude, J_polita$latitude), col="white", pch=20, cex=.8)
```
          
Let's build and combine your new skills in looking up species occurrence data with what we learned about mapping. 

Going back to our peninsula, here are a bunch of species that are endemic to that area.

Spergularia melanocaulos
Petrocoptis grandiflora
Spergula viscosa
Minuartia cerastiifolia
Minuartia recurva
Petrocoptis hispanica
Petrocoptis pyrenaica

Choose any two, and do the following:
```{r} 
#1. Look up - are they accepted names or are they synonyms?
library("rgbif")
library("spocc")
library("taxize")
#Sometimes you know the species name, but you may be a little unsure of the spelling 
results=gnr_resolve(names="Malus Pumila")
head(results$matched_name)
results=gnr_resolve(names="Solanum melongena")
head(results$matched_name)
#this is the accepted name for apple tree

```

```{r}
name="Malus Pumila"  
tsn=get_tsn(name, accepted=FALSE)
tsn
#The output suggests checking the itis database for the accepted name 
lapply(tsn, itis_acceptname)
#now we have the official taxon number associated with apples
```

```{r}
#2. Look up the two plant taxa. Are they both angiosperms?
#3. Angiosperm or otherwise, what plant family do they belong to?
classification(c("Malus pumila"), db="itis")
classification(c("Solanum melongena"), db="itis")
#yes they are angiosperms
```
```{r}
#4. For at least one genus, look up all the species in that genus.
downstream("Solanum", downto="species", db="itis") 
#here we see that there are many species in this genus
```

```{r}
#5. For one species, look up to see if there are subspecies or other
children("Malus pumila", db="col")
#this shows us that there are  several infraspecies
```

```{r}
#6. For both, search and download species occurrence data; prepare as a data-frame
library(rgbif)
pumila <- occ_search(scientificName="Malus pumila", limit=1000)
#head(pumila)
colnames(pumila$data)
#pumila$data$genus
#pumila$data$species
unique(pumila$data$genus)
unique(pumila$data$species)
melongena <- occ_search(scientificName="Solanum melongena", limit=1000)
#head(melongena)
colnames(melongena$data)
#melongena$data$genus
#melongena$data$species
unique(melongena$data$genus)
unique(melongena$data$species)
```
```{r}
#7. Clean-up the data-frame:remove occurrences with missing data and duplicates
#Choose just three variables
pumila.small <- subset(pumila$data, select=c("species", "decimalLatitude",
"decimalLongitude"))
#Rename the columns, make sure you're working with a dataframe
colnames(pumila.small) <- c("species", "lat", "lon")
pumila.small <- as.data.frame(pumila.small)
head(pumila.small)
tail(pumila.small)
#There are often occurrences with missing values to remove 
dim(pumila.small)
pumila.small <- pumila.small[complete.cases(pumila.small),]
dim(pumila.small)
#Also there may be duplicates, the same plant or population observed twice
#Or uploaded twice into the database
pumila.u <- unique(pumila.small)
dim(pumila.u)
```

9. Also map that species with climate variables on a world map or European map. 
10. Returning to your results for #4 above, find another species with a range outside of the  peninsula. Download occurrence data and map on a world map. 
11. Then re-map on an appropriately cropped map.

ONE LAST THING
12. Don't forget to identify the peninsula that we are working with! 
```{r}
#WORK WITH OCCURRENCE DATA, WORLDCLIM DATA AND MAPS
#8. Map occurrence data with a climate variable for one species
#We can get occurrence 
appletree <- occ(q="Malus pumila", from="gbif", limit=100)
appletree
#getting data
appletree.dat <- (appletree$gbif$data)
head(appletree.dat)
#slimming it down
appletree.df <- occ2df(DawnRedwood)
appletree.df
#plotting occurence data on map with altitude
plot(all.worldalt[["alt"]])
points(appletree.df[,c("longitude", "latitude")], pch = 16, cex = 0.5, col = "red")
#9 mapping with another climate variable with zoomed in range
plot(all.worldalt[["alt"]], xlim=c(-25,50),ylim=c(35,71))
points(appletree.df[,c("longitude", "latitude")], pch = 16, cex = 0.5, col = "red")
```
```{r}
#10 and 11: doing the same thing with another species 
melongena.small <- subset(melongena$data, select=c("species", "decimalLatitude",
"decimalLongitude"))
colnames(melongena.small) <- c("species", "lat", "lon")
melongena.small <- as.data.frame(melongena.small)
head(melongena.small)
tail(melongena.small)
#There are often occurrences with missing values to remove 
dim(melongena.small)
melongena.small <- melongena.small[complete.cases(melongena.small),]
dim(melongena.small)
melongena.u <- unique(melongena.small)
dim(melongena.u)
```
```{r}
#plotting eggplant
eggplant <- occ(q="Solanum melongena", from="gbif", limit=100)
eggplant
#getting data
eggplant.dat <- (eggplant$gbif$data)
head(eggplant.dat)
#slimming it down
eggplant.df <- occ2df(eggplant)
eggplant.df
#plotting occurence data on map with altitude
plot(all.worldalt[["alt"]])
points(eggplant.df[,c("longitude", "latitude")], pch = 16, cex = 0.5, col = "red")
#9 mapping with another climate variable with zoomed in range
plot(all.worldalt[["alt"]], xlim=c(-25,50),ylim=c(35,71))
points(eggplant.df[,c("longitude", "latitude")], pch = 16, cex = 0.5, col = "red")
```

